---
title: "feventr quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{feventr quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
```

`feventr` estimates causal effects of events on financial returns. Every
estimator consumes the same long panel — unit, date, return, plus treatment
assignment and an event date — and returns a common `fes_fit` object.

## Simulated data

The built-in DGP is the two-factor structure from Goldsmith-Pinkham & Lyu's
simulation section: factor loadings drawn around one, daily Fama-French factor
realizations, idiosyncratic noise calibrated so half the return variance is
systematic, and a one-shot treatment effect on the event day. Selection
mechanisms (`"assignment"`, `"timing"`, `"both"`) reproduce the paper's
selection-on-loadings and selection-on-factor-timing designs.

```{r}
library(feventr)
sim <- simulate_events(n_units = 200, n_pre = 120, n_candidate = 1,
                       n_post = 10, tau = 0.03, seed = 42)
str(sim$events)
```

## Fitting estimators

One entry point, `event_study()`; the estimator is an argument. `returns` is
required — it records whether `ret` holds simple or log returns and selects
how cumulative effects compound.

```{r}
args <- list(data = sim$data, unit = "id", time = "t", ret = "ret",
             treated = sim$events$unit, event_time = sim$event_time,
             window = c(0, 10), est_window = c(-120, -1), returns = "simple")

fit_mean <- do.call(event_study, c(args, list(method = "mean")))
fit_sc   <- do.call(event_study, c(args, list(method = "sc", se = "placebo",
                                              reps = 50, seed = 1)))
fit_sc
```

The event-day ATT recovers the simulated 3% effect. `summary()` gives the
per-period path; `coef()`, `vcov()`, and `confint()` behave as usual.

```{r}
head(summary(fit_sc))
```

Factor-model abnormal returns take the factor series as a data frame keyed on
the panel's time column (excess-return adjustments are the caller's choice —
CAPM means passing excess returns and Mkt-RF):

```{r}
fit_2f <- do.call(event_study, c(args, list(
  method = "factor", factors = sim$factors[, c("t", "mktrf", "smb")])))
round(fit_2f$att, 4)
```

## Plots and diagnostics

```{r}
plot(fit_sc)
plot(fit_sc, what = "paths")
```

`event_betas()` compares treated, control, and synthetic-control-weighted
factor loadings — the paper's core diagnostic for when simple comparisons
mislead:

```{r}
event_betas(fit_sc, sim$factors[, c("t", "mktrf", "smb")])
```

## Many events at once

`event_study_batch()` fits one model per event (each with its own event date,
treated units, and donor pool, excluding contaminated donors), in parallel,
and reports cross-event averages with cross-event standard errors:

```{r}
ev <- data.frame(unit = c("1", "2", "10"), event_time = c(80, 80, 100),
                 event = c("a", "a", "b"))
b <- event_study_batch(sim$data, "id", "t", "ret", events = ev,
                       method = "sc", window = c(0, 5),
                       est_window = c(-60, -1), returns = "simple")
b
summary(b)
```

## The solver

Synthetic control with thousands of donors is the package's reason to exist:
`solve_simplex_ls()` runs Frank-Wolfe (never forming the donor Gram matrix) to
find the sparse active donor set, polishes with an exact QP on that support,
and verifies optimality with a KKT gradient screen. See the README benchmark
table; the solver is exported for direct use:

```{r}
A <- matrix(rnorm(50 * 200), 50, 200)   # 200 donors, 50 pre-periods
w_true <- c(rep(0.25, 4), rep(0, 196))
b_ <- as.vector(A %*% w_true) + rnorm(50, sd = 0.05)
sol <- solve_simplex_ls(A, b_)
round(head(sort(sol$w, decreasing = TRUE), 6), 3)
```
